\section{Summary}

This thesis set out to build skin-lesion classifiers for ISIC-2019 that are both
accurate across all eight classes and practical to deploy, and that use the
clinical metadata a dermatologist would normally have on hand. We approached the
problem with a single design philosophy --- convolutional features for local
texture, a transformer trunk for global structure, and a cross-attention head
that folds in age, sex, and lesion site --- realised at two points on the
efficiency curve. The lightweight hybrid, at 3.98M parameters and 0.631 GMACs,
reaches a balanced multi-class accuracy of 0.6081 and is the most
parameter-efficient model in the sub-6M budget, beating all five lightweight
baselines including the strong hybrid baselines MobileViT-S and
EfficientFormer-L1. The flagship DEKAN model, at 16.45M parameters, reaches
0.6438 BMA, the best in the study and comparable to the balanced accuracy of the
top ISIC-2019 leaderboard ensemble, in a single model.

\section{Contributions}

The main contributions of this work are: (1) a lightweight hybrid
CNN--Transformer with clinical-metadata cross-attention that handles missing
fields with learned embeddings and meets an explicit budget of under 6M
parameters and 1 GMAC; (2) DEKAN, a higher-capacity model in the same family
that fuses two complementary pretrained backbones through learned attention and
adds a Kolmogorov--Arnold classifier head; (3) a fair, single-protocol
comparison of both models against five established baselines, with controlled
ablations that quantify what each component adds; and (4) two honest negative or
inconclusive results --- that the KAN head did not beat a linear head on this
task, and that the metadata benefit for the lightweight model could not be
confirmed on a single test evaluation.

\section{Future Work}

The most immediate next steps follow directly from the limitations in
Chapter~5. Repeating every run across at least three seeds and reporting mean and
standard deviation would put the comparisons on a publication-grade footing.
Completing the three unfinished ablations --- the pure-ViT lightweight variant
and the two single-backbone DEKAN variants --- would isolate the value of the
CNN inductive bias and pin down the size of the dual-backbone fusion gain.
Evaluating on an external dataset, for example mapping the eight-class output to
a melanoma-versus-rest decision on a separate collection, would test
generalisation beyond ISIC-2019. Finally, qualitative interpretability ---
Grad-CAM on the CNN stem and attention rollout on the transformer trunk --- would
help confirm that the models attend to clinically meaningful regions, which
matters for any tool intended to support, rather than replace, a clinician.
